---
title: "Cell Ranger QC report"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
params:
  precomputed: !r NULL
  include_h5: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(cellrangercheck)
pc <- params$precomputed
long <- pc$long; meta <- pc$meta; wide <- pc$wide
flags <- pc$flags; status <- pc$status; br <- pc$br; dist <- pc$dist
fqc <- pc$fqc; mdat <- pc$mdat
```

# Run overview

Number of GEM wells: **`r nrow(meta)`**. Pipelines:
`r paste(sort(unique(long$pipeline)), collapse = ", ")`.

```{r meta}
DT::datatable(meta, rownames = FALSE, options = list(dom = "t"))
```

# Overall QC status

```{r status}
cap <- if ("fastqc_status" %in% names(status)) {
  "overall = worst of the Cell Ranger metric status and the FastQC status"
} else {
  "overall = Cell Ranger metric status"
}
DT::datatable(status, rownames = FALSE, caption = cap, options = list(dom = "t"))
```

# Comparable per-sample metrics

```{r wide}
DT::datatable(wide, rownames = FALSE,
              options = list(scrollX = TRUE, pageLength = 25))
```

# Metric explorer {.tabset}

Pick a metric from the dropdown to plot it across libraries (x = GEM well). One
tab per assay.

```{r metric_explorer, results='asis'}
if (requireNamespace("plotly", quietly = TRUE) && !is.null(mdat) && nrow(mdat)) {
  for (a in sort(unique(mdat$library_type))) {
    src <- c(
      sprintf("## %s", a), "",
      "```{r, echo=FALSE}",
      sprintf("assay_metric_plot(mdat, %s)", deparse(a)),
      "```", ""
    )
    cat(knitr::knit_child(text = src, envir = environment(), quiet = TRUE))
  }
} else {
  cat("Install the 'plotly' package to enable the interactive metric explorer.\n")
}
```

# Flagged metrics

```{r flags}
library(ggplot2)
if (nrow(flags)) {
  flags$label <- paste(flags$library_type, flags$metric, sep = ": ")
  ggplot(flags, aes(x = sample_id, y = label, fill = status)) +
    geom_tile(color = "white") +
    scale_fill_manual(values = c(pass = "#2c7fb8", warn = "#fec44f",
                                 fail = "#de2d26"), na.value = "grey85") +
    labs(x = NULL, y = NULL, fill = "status") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
} else {
  cat("No metrics matched the supplied thresholds.")
}
```

```{r flags_table}
DT::datatable(flags, rownames = FALSE, options = list(scrollX = TRUE))
```

# Barcode-rank curves

```{r barcode_ranks, fig.width=7, fig.height=4}
if (isTRUE(params$include_h5) && !is.null(br) && nrow(br)) {
  ggplot(br, aes(rank, total_umi, color = run_id)) +
    geom_line() +
    geom_hline(aes(yintercept = knee, color = run_id), linetype = "dashed") +
    scale_x_log10() + scale_y_log10() +
    labs(x = "Barcode rank", y = "Total UMI", color = "GEM well") +
    theme_minimal()
}
```

# Per-cell UMI / gene distributions

```{r umi_gene}
if (isTRUE(params$include_h5) && !is.null(dist)) {
  DT::datatable(dist, rownames = FALSE, options = list(scrollX = TRUE, dom = "t"))
}
```

```{r fastqc_section, results='asis', eval=!is.null(fqc)}
cat("# FastQC\n")
```

```{r fastqc_stats, eval=!is.null(fqc)}
fs <- fqc$stats
fs <- fs[order(fs$run_id, fs$fastq_id, fs$read), ]
show <- fs[, c("run_id", "fastq_id", "feature_types", "read", "filename",
               "total_sequences", "read_length", "pct_gc", "pct_dup")]
DT::datatable(show, rownames = FALSE, caption = "Basic statistics per fastq file",
              options = list(scrollX = TRUE, pageLength = 25))
```

```{r fastqc_modules, fig.width=9, fig.height=5, eval=!is.null(fqc)}
fm <- fqc$modules
if (!is.null(fm) && nrow(fm)) {
  fm$file_lab <- paste(ifelse(is.na(fm$run_id), "?", fm$run_id),
                       fm$fastq_id, fm$read, sep = " | ")
  fm$module <- factor(fm$module, levels = rev(unique(fm$module)))
  fm$status <- factor(fm$status, levels = c("PASS", "WARN", "FAIL"))
  ggplot(fm, aes(x = file_lab, y = module, fill = status)) +
    geom_tile(color = "white") +
    scale_fill_manual(values = c(PASS = "#2c7fb8", WARN = "#fec44f",
                                 FAIL = "#de2d26"), na.value = "grey85") +
    labs(x = NULL, y = NULL, fill = "FastQC",
         title = "FastQC module status by reaction | library | read") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

```{r fastqc_flagged_hdr, results='asis', eval=!is.null(fqc)}
cat("\n## Flagged modules\n\nModules that did not PASS — the detail behind the heatmap.\n")
```

```{r fastqc_flagged, eval=!is.null(fqc)}
fm <- fqc$modules
fl <- fm[fm$status %in% c("WARN", "FAIL"),
         c("run_id", "fastq_id", "read", "module", "status")]
fl <- fl[order(fl$status != "FAIL", fl$run_id, fl$fastq_id), ]
DT::datatable(fl, rownames = FALSE, options = list(scrollX = TRUE, pageLength = 15))
```

```{r fastqc_overrep_hdr, results='asis', eval=!is.null(fqc)}
cat("\n## Overrepresented sequences\n")
```

```{r fastqc_overrep, eval=!is.null(fqc)}
ov <- fqc$overrep
if (!is.null(ov) && nrow(ov)) {
  cols <- intersect(c("run_id", "fastq_id", "read", "sequence", "count",
                      "percentage", "possible_source"), names(ov))
  ov <- ov[order(-ov$percentage), cols]
  DT::datatable(ov, rownames = FALSE,
                caption = "Overrepresented sequences reported by FastQC",
                options = list(scrollX = TRUE, pageLength = 15))
} else {
  cat("No overrepresented sequences reported.")
}
```
